\Large\bf FlexibleSUSY 2.0: Extensions to investigate the phenomenology of SUSY and non-SUSY models

### 2018-04-28 doc/flexiblesusy-2.0.tex
changes made by CPC

The resulting electron EDM is displayed as contours in
\figref{fig:EDM plots}.
In the left panel, $\phi_{\mhalf} \equiv \arg \mhalf$ and
$\phi_\mu \equiv \arg \mu$ are varied
while $\arg(\mu A_e)$ is constrained to be zero.
From the directions of the contours it is clear
that \fs correctly reproduces the behaviour of
$\edm{e}$ depending only on the rephasing invariant $\arg(\mu \mhalf)$.
In the right panel, the roles of $\phi_{A_e} \equiv \arg A_e$ and
$\phi_{\mhalf}$ are swapped so that $\arg(\mu \mhalf)$ stays at zero.
Again, the contours verify that
$\edm{e}$ from \fs is determined by the physical phase $\arg(\mu A_e)$.
For reference,
the experimental upper limit on $|\edm{e}|$ at the 90\% confidence level
is shown as the thick red lines \cite{Baron:2013eja}.

\subsection{Tests and comparisons with other spectrum generators}
%
In the left panel of \figref{fig:CMSSMCPV_plots}, we show the lightest
Higgs pole mass calculated in the \CP-violating CMSSM at the 1-loop
level with \fstwo and \spheno~4.0.2 as a function of the phase angle
of the complex $\mu$ parameter, $\phi_\mu \equiv \arg \mu$.
We use a low-energy scenario with $\mzero = \mhalf = 500\unit{GeV}$,
$\tan\beta = 10$ and $\azero = 0$.  Even though this scenario is
excluded, the figure illustrates that both \fs and \spheno show the
same behaviour of the Higgs pole mass as a function of $\phi_\mu$.
The $\phi_\mu$-independent shift of around $0.7\unit{GeV}$ between the
Higgs masses calculated by the two programs is mainly caused by the
different treatment of higher-order corrections to
the running \DRbar top Yukawa coupling, which has been discussed in
Refs.~\cite{Staub:2015aea,ThomasKwasnitza:2016yqj}.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_Mh_PhiMu_m0-500_M12-500_TB-10_A0-0}\hfill
  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_TYu33_Q_m0-500_M12-500_TB-10_A0-500_PhiMu-0_PhiA0-Pi4}
  \caption{Left panel: lightest Higgs pole mass calculated in the
    \CP-violating CMSSM at the 1-loop level with \fstwo and
    \spheno 4.0.2 as a function of the phase angle of the $\mu$
    parameter for $\mzero = \mhalf = 500\unit{GeV}$, $\tan\beta = 10$
    and $\azero = 0$.  Right panel: renormalization group running of the
    trilinear coupling $(T^u)_{33}$ for $\mzero = \mhalf =
    500\unit{GeV}$, $\tan\beta = 10$, $\phi_\mu = 0$ and $\azero = 500\,
    e^{i \pi/4} \unit{GeV}$.}
  \label{fig:CMSSMCPV_plots}
\end{figure}

@@

behaviour here, it is necessary to either fine-tune the provided input

@@

The first potential source of large 2-loop logarithms in the matching
procedure is the insertion of parameters, which have been obtained from the SM
via a 1-loop matching, into the 1-loop self energies or tadpoles of the \BSM
model.  We illustrate this effect with the most important parameter, the top
Yukawa coupling:
The running top Yukawa coupling of the \BSM model  $y_t^\BSM$ is
determined by a matching as\footnote{We ignore potential tree-level
  factors here for brevity.}
%
\begin{align}
  y_t^\BSM = y_t^\SM + \Delta y_t \,, \label{eq:yt_SM_BSM}
\end{align}
%
where $\Delta y_t$ is of 1-loop order (but without large logarithms).
At the same time, the Higgs pole mass calculations on the left-hand side and
right-hand side of Eq.~\eqref{eq:Higgs_pole_mass_matching} are of the form
%
\begin{align}
  (M_h^\SM)^2 &= (m_h^\SM)^2 + \propto \frac{(v^\SM)^2(y_t^\SM)^4}{(4\pi)^2} \log\frac{m_t^\SM}{Q} + \cdots \,,
  \label{eq:Mh_top_SM}\\
  (M_h^\BSM)^2 &= (m_h^\BSM)^2 + \propto \frac{(v^\BSM)^2(y_t^\BSM)^4}{(4\pi)^2} \log\frac{m_t^\BSM}{Q} + \cdots \,,
  \label{eq:Mh_top_BSM}
\end{align}
%
where the matching scale $Q$ is of the order $\MS$ and we have also
introduced $m_h^\SM$, $m_t^\SM$ and $m_t^\BSM$ for the running SM
Higgs mass, running SM top mass and running BSM top mass,
keeping our convention of using an upper case `$M$' for pole masses
and lower case `$m$' for running tree-level masses.  If
Eq.\ \eqref{eq:Mh_top_SM} and Eq.\ \eqref{eq:Mh_top_BSM} are set equal
and the relation Eq.\ \eqref{eq:yt_SM_BSM} is inserted, potentially
large 2-loop terms for example of the form
%
\begin{align}
  \frac{4(v^\SM)^2(y_t^\SM)^3 \Delta y_t}{(4\pi)^2} \log\frac{m_t^\SM}{Q} + \cdots
\label{eq:twolooplogsyt}
\end{align}
%
remain. Such terms effectively shift the quartic Higgs coupling of the
SM by  next-to-leading logarithmic 2-loop terms.%
\footnote{If the self energies are evaluated at the
  2-loop level, the problem repeats itself one order higher, i.e., the
  term in Eq.~\eqref{eq:twolooplogsyt} is cancelled but similar terms of
  next-to-leading logarithmic 3-loop order remain.}

>>>>

The resulting electron EDM is displayed as contours in
\figref{fig:EDM plots}.
In the left panel, $\phi_{\mhalf} \equiv \arg \mhalf$ and
$\phi_\mu \equiv \arg \mu$ are varied
while $\arg(\mu A_e)$ is constrained to be zero.
From the directions of the contours it is clear
that \fs correctly reproduces the behavior of
$\edm{e}$ depending only on the rephasing invariant $\arg(\mu \mhalf)$.
In the right panel, the roles of $\phi_{A_e} \equiv \arg A_e$ and
$\phi_{\mhalf}$ are swapped so that $\arg(\mu \mhalf)$ stays at zero.
Again, the contours verify that
$\edm{e}$ from \fs is determined by the physical phase $\arg(\mu A_e)$.
For reference,
the experimental upper limit on $|\edm{e}|$ at the 90\% confidence level
is shown as the thick red lines \cite{Baron:2013eja}.

\subsection{Tests and comparisons with other spectrum generators}
%
In the left panel of \figref{fig:CMSSMCPV_plots}, we show the lightest
Higgs pole mass calculated in the \CP-violating CMSSM at the 1-loop
level with \fstwo and \spheno~4.0.2 as a function of the phase angle
of the complex $\mu$ parameter, $\phi_\mu \equiv \arg \mu$.
We use a low-energy scenario with $\mzero = \mhalf = 500\unit{GeV}$,
$\tan\beta = 10$ and $\azero = 0$.  Even though this scenario is
excluded, the figure illustrates that both \fs and \spheno show the
same behavior of the Higgs pole mass as a function of $\phi_\mu$.
The $\phi_\mu$-independent shift of around $0.7\unit{GeV}$ between the
Higgs masses calculated by the two programs is mainly caused by the
different treatment of higher-order corrections to
the running \DRbar top Yukawa coupling, which has been discussed in
Refs.~\cite{Staub:2015aea,ThomasKwasnitza:2016yqj}.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_Mh_PhiMu_m0-500_M12-500_TB-10_A0-0}\hfill
  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_TYu33_Q_m0-500_M12-500_TB-10_A0-500_PhiMu-0_PhiA0-Pi4}
  \caption{Left panel: lightest Higgs pole mass calculated in the
    \CP-violating CMSSM at the 1-loop level with \fstwo and
    \spheno 4.0.2 as a function of the phase angle of the $\mu$
    parameter for $\mzero = \mhalf = 500\unit{GeV}$, $\tan\beta = 10$
    and $\azero = 0$.  Right panel: renormalization group running of the
    trilinear coupling $(T^u)_{33}$ for $\mzero = \mhalf =
    500\unit{GeV}$, $\tan\beta = 10$, $\phi_\mu = 0$ and $\azero = 500\,
    e^{i \pi/4} \unit{GeV}$.}
  \label{fig:CMSSMCPV_plots}
\end{figure}

@@

behavior here, it is necessary to either fine-tune the provided input

@@

The first potential source of large 2-loop logarithms in the matching
procedure is the insertion of parameters, which have been obtained from the SM
via a 1-loop matching, into the 1-loop self energies or tadpoles of the \BSM
model.  We illustrate this effect with the most important parameter, the top
Yukawa coupling:
The running top Yukawa coupling of the \BSM model  $y_t^\BSM$ is
determined by a matching as\footnote{We ignore potential tree-level
  factors here for brevity.}
%
\begin{align}
  y_t^\BSM = y_t^\SM + \Delta y_t \,, \label{eq:yt_SM_BSM}
\end{align}
%
where $\Delta y_t$ is of 1-loop order (but without large logarithms).
At the same time, the Higgs pole mass calculations on the left-hand side and
right-hand side of Eq.~\eqref{eq:Higgs_pole_mass_matching} are of the form
%
\begin{align}
  (M_h^\SM)^2 &= (m_h^\SM)^2 + \propto \frac{(v^\SM)^2(y_t^\SM)^4}{(4\pi)^2} \log\frac{m_t^\SM}{Q} + \cdots \,,
  \label{eq:Mh_top_SM}\\
  (M_h^\BSM)^2 &= (m_h^\BSM)^2 + \propto \frac{(v^\BSM)^2(y_t^\BSM)^4}{(4\pi)^2} \log\frac{m_t^\BSM}{Q} + \cdots \,,
  \label{eq:Mh_top_BSM}
\end{align}
%
where the matching scale $Q$ is of the order $\MS$ and we have also
introduced $m_h^\SM$, $m_t^\SM$ and $m_t^\BSM$ for the running SM
Higgs mass, running SM top mass and running BSM top mass,
keeping our convention of using an upper case `$M$' for pole masses
and lower case `$m$' for running tree-level masses.  If
Eq.\ \eqref{eq:Mh_top_SM} and Eq.\ \eqref{eq:Mh_top_BSM} are set equal
and the relation Eq.\ \eqref{eq:yt_SM_BSM} is inserted, potentially
large 2-loop terms for example of the form
%
\begin{align}
  \frac{4(v^\SM)^2(y_t^\SM)^3 \Delta y_t}{(4\pi)^2} \log\frac{m_t^\SM}{Q} + \cdots
\label{eq:twolooplogsyt}
\end{align}
%
remain. Such terms effectively shift the quartic Higgs coupling of the
SM by  next-to-leading logarithmic 2-loop terms.%
\footnote{If the self energies are evaluated at the
  2-loop level, the problem repeats itself one order higher, i.e., the
  term in Eq.~\eqref{eq:twolooplogsyt} is canceled but similar terms of
  next-to-leading logarithmic 3-loop order remain.}
